A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-01-29 src/tex/ms.tex
Updates from Overleaf

The two-infall model has long been proposed to explain the formation of the Milky Way's thick and thin disks. The infall history of any given region of the disk is described by two successive, exponentially declining bursts, whose timescale may vary as a function of radius. The first infall, which forms the thick disk, has a short timescale, typically on the order of $\sim0.1-1$ Gyr. The second infall forms the thin disk and has a much longer timescale, typically $\sim4-10$ Gyr.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models (see discussion in Section \ref{sec:methods}). We omit some parameters that are unchanged from \citetalias{johnson_stellar_2021}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Disk lifetime \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{johnson_impact_2020} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & Delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \tau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \tau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst infall rate \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall infall rate \\
        \hline
        % One-zone parameters
        \tau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \tau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm on})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm on}$ is the onset time of the second infall epoch, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_{\Sigma,0} e^{R(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0$ kpc, a thin disk scale radius of $R_2=2.5$ kpc, and a central surface density ratio of $f_{\Sigma,0}=0.27$ \citet{bland-hawthorn_galaxy_2016}.

\begin{figure*}
    \centering
    \includegraphics{figures/onezone_params.pdf}
    \caption{Gas abundance tracks in the [O/Fe]--[Fe/H] plane for one-zone chemical evolution models which assume different values for the infall history parameters. In each panel, one parameter is varied according to the legend while the other two are held fixed. The open symbols along each curve mark logarithmic steps in time, as denoted in panel (b). The marginal panels show the corresponding stellar abundance distributions, which are convolved with a Gaussian kernel with a width of 0.02 dex for visual clarity. All models use the {\tt yZ1} yield set and assume $\eta=0.4$. {\bf (a)} The first infall timescale $\tau_1$ primarily affects the stellar distribution along the high-$\alpha$ sequence. {\bf (b)} The second infall timescale $\tau_2$ affects the [O/Fe] distribution of the low-$\alpha$ sequence and the width of the MDF. {\bf (c)} The onset time $t_{\rm on}$ has a minor effect on the position of the abundance distributions for values $t_{\rm on}\ge 3$ Gyr. While the ISM abundance at $t_{\rm on}$ varies widely for some parameter choices, it has a negligible impact on the stellar abundance distributions. {\bf Key takeaway:} The second infall timescale $\tau_2$ is the most important parameter to fit the low-$\alpha$ sequence and MDF.}
    \label{fig:twoinfall-parameters-low-yields}
\end{figure*}

@@

@@

@@

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_df_yields.pdf}
    \caption{Stellar [O/Fe] distributions produced by multi-zone models (a, b) and as observed by APOGEE (c). Stars are binned by Galactocentric radius, represented by the color scale, and absolute midplane distance $|z|$, represented by the different rows. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. {\bf (a)} Multi-zone model output assuming the {\tt yZ1} yield set. {\bf (b)} Model output assuming the {\tt yZ2} yield set. {\bf (c)} The distributions from APOGEE DR17 for reference. {\bf Key takeaway:} }
    \label{fig:ofe-df-yields}
\end{figure*}

>>>>

The two-infall model has long been proposed to explain the formation of the Milky Way's thick and thin disks. The infall history of any given region of the disk is described by two successive, exponentially declining bursts, which explain the origin of the thick and thin disks respectively.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of variables and their fiducial values for our chemical evolution models (see discussion in Section \ref{sec:methods}).\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        \sigma_{\rm RM8}    & 2.68 kpc  & \ref{sec:migration}   & Radial migration strength \\
        % Outflows
        R_\eta          & 5.4 kpc   & \ref{sec:outflows}    & Outflow scale radius \\
        % Yields
        y/Z_\odot       & 1         & \ref{sec:yields}  & Nucleosynthetic yield scaling (see Table \ref{tab:yields}) \\
        f_{\rm Ia}(t)   & Plateau   & \ref{sec:yields}  & Delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:yields}  & Minimum SN Ia delay time \\
        % Star formation histories
        % f_{\rm in}(t|R_{\rm gal})   & Equation \ref{eq:infall-rate}   & \ref{sec:sfh} & Time-dependence of the gas infall rate \\
        \tau_1          & 1 Gyr     & \ref{sec:sfh}     & Timescale of the first infall epoch \\
        \tau_2          & 15 Gyr    & \ref{sec:sfh}     & Timescale of the second infall epoch \\
        t_{\rm on}      & 4.2       & \ref{sec:sfh}     & Onset time of the second infall epoch \\
        f_{\Sigma,0}    & 0.27      & \ref{sec:sfh}     & Central thick/thin disk surface density ratio \\
        {\rm [X/H]}_{\rm CGM}   & Pristine  &           & Pre-enrichment of infalling gas
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

\begin{equation}
    \eta(R_{\rm gal}) = \eta_\odot \exp\Big(\frac{R_{\rm gal}-R_\odot}{R_\eta}\Big)
\end{equation}

\begin{equation}
    \eta_\odot = \frac{y_{\rm O}^{\rm CC}}{Z_{\rm O,\odot}} - 1 + r
\end{equation}

\begin{equation}
    R_\eta = -\frac{1}{\nabla{\rm [O/H]}_{\rm eq} \times\ln(10)}
\end{equation}

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm on})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm on}$ is the onset time of the second infall epoch, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_{\Sigma,0} e^{R(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0$ kpc, a thin disk scale radius of $R_2=2.5$ kpc, and a central surface density ratio of $f_{\Sigma,0}=0.27$ \citep{bland-hawthorn_galaxy_2016}.

The parameters of Equation \ref{eq:infall-rate} have an important effect on the model output. Previous studies of the two-infall model have adopted values for the infall parameters of $\tau_1\sim0.1-1$ Gyr, $\tau_2\sim4-10$ Gyr, and $t_{\rm on}\sim4$ Gyr. Figure \ref{fig:twoinfall-parameters} illustrates the effect of varying each of these parameters on gas abundance tracks and stellar abundance distributions in a one-zone model. The first infall timescale $\tau_1$ primarily affects the stellar distribution along the high-$\alpha$ sequence, with negligible effect on the low-$\alpha$ sequence (a). On the other hand, the second infall timescale $\tau_2$ controls the width of the MDF and the [O/Fe] distribution of the low-$\alpha$ sequence, with a shorter $\tau_2$ producing a distribution which is skewed to higher [O/Fe] (b). The onset time $t_{\rm on}$ strongly affects the overall stellar abundance distribution for values $t_{\rm on}\leq2$ Gyr, but in this case the gas tracks do not produce the characteristic loop characteristic of the two-infall model. For $t_{\rm on}>2$ Gyr, varying $t_{\rm on}$ results in a minor shift to the mean of the MDF and little change to the [O/Fe] distributions, even though the abundance tracks in [O/Fe]--[Fe/H] space appear very different. We find that $\tau_2$ is the most important parameter to tune in order to reproduce the observed stellar abundance distributions.

\begin{figure*}
    \centering
    \includegraphics{figures/onezone_params.pdf}
    \caption{Gas abundance tracks in the [O/Fe]--[Fe/H] plane for one-zone chemical evolution models which assume different values for the infall history parameters. In each panel, one parameter is varied according to the legend while the other two are held fixed. The open symbols along each curve mark logarithmic steps in time, as denoted in panel (b). The marginal panels show the corresponding stellar abundance distributions, which are convolved with a Gaussian kernel with a width of 0.02 dex for visual clarity. All models use the {\tt yZ1} yield set and assume $\eta=0.4$.}
    \label{fig:twoinfall-parameters}
\end{figure*}

@@

\subsection{Stellar Migration}
\label{sec:migration}

\begin{equation}
    \sigma_{\rm RM} = \sigma_{\rm RM8} \Big(\frac{\tau}{8\,{\rm Gyr}}\Big)^{0.33} \Big(\frac{R_{\rm form}}{8\,{\rm kpc}}\Big)^{0.61}
    \label{eq:radial-migration}
\end{equation}

\begin{equation}
    p(z|\tau,R_{\rm final}) = \frac{1}{4 h_z} {\rm sech}^2\Big(\frac{z_{\rm final}}{2 h_z(\tau,R_{\rm final})}\Big).
    \label{eq:sech-pdf}
\end{equation}

\begin{equation}
    h_z(\tau,R_{\rm final}) = (h_{z,s}/e^2) \exp(\tau/\tau_s + R_{\rm final}/R_s)
    \label{eq:scale-height}
\end{equation}
where $h_{z,s}=0.24$ kpc is the scale height at $\tau_s=7$ Gyr and $R_s=6$ kpc.

@@

\subsection{Stellar Abundance Distributions}

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_df_yields.pdf}
    \caption{Stellar [O/Fe] distributions produced by multi-zone models (a--d) and as observed by APOGEE (d). Stars are binned by Galactocentric radius, represented by the color scale, and absolute midplane distance $|z|$, represented by the different rows. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties were forward-modeled onto the model outputs. For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. Model (a) assumes the fiducial parameters and the {\tt yZ1} yield set, whereas models (b--d) assume the {\tt yZ2} yield set. Model (c) also assumes a power-law SN Ia DTD, while model (d) assumes the infalling gas has metallicity ${\rm [O/H]}_{\rm CGM}={\rm [Fe/H]}_{\rm CGM}=-0.7$. The distributions from APOGEE DR17 are plotted in column (e) for reference. {\bf Key takeaway:} The high-yield case produces a peak at intermediate [O/Fe] which is prominent in most galactic regions, which can be mitigated by pre-enriched infall.}
    \label{fig:ofe-df-yields}
\end{figure*}

\subsection{Age--Abundance Relations}

@@
